Wald test to evaluate regression and dispersion parameters in multivariate covariance generalized linear models

### 2022-07-11 5_artigos/statistic-in-medicine/draft/DRAFT1/draft-sm-lineu.tex
atualiza artigos

The power variance function characterizes the Tweedie family of distributions, it is given by $\vartheta\left(\cdot; p_r\right) = \mu^{p_r}_r$, in which some distributions stand out: Normal ($p $ = 0), Poisson ($p$ = 1), gamma ($p$ = 2) and Inverse Normal ($p$ = 3) \cite{Jorgensen87, Jorgensen97}.

The Poisson–Tweedie dispersion function \cite{Jorgensen15} is an option for responses that characterize counts. The dispersion function is given by $\vartheta\left(\cdot; p\right) = \mu + \tau\mu^p$ where $\tau$ is the dispersion parameter. We thus have a rich class of models to deal with responses that characterize counts, since many important distributions appear as special cases, such as: Hermite ($p$ = 0), Neyman type A ($p$ = 1), binomial negative ($p$ = 2) and Poisson–Gaussian inverse (p = $3$).

@@

The dispersion matrix $\boldsymbol{\Omega({\tau})}$ describes the part of the covariance within each response variable that does not depend on the mean structure, ie, the correlation structure between the sample observations. Based on the ideas of \cite{Anderson73} and \cite{Pourahmadi00}, \cite{Bonat16} proposed to model the dispersion matrix through a linear matrix predictor combined with a covariance link function given by:

@@

In this way, the McGLMs configure a general framework for analysis via regression models for non-Gaussian data with multiple responses in which no assumptions are made regarding the independence of the observations. The class is defined by 3 functions (link, variance and covariance) in addition to a linear predictor and a linear matrix predictor for each response under analysis.

@@

McGLMs are fitted based on the method of estimation functions described in detail by \cite{Bonat16} and \cite{jorg04}. This subsection presents an overview of the algorithm and the asymptotic distribution of estimators based on estimation functions.

@@

For the dispersion parameters, the Pearson estimation function is used, defined in the form

@@

\noindent where $W_{\boldsymbol{\lambda}i} = -\frac{\partial \boldsymbol{C}^{-1}}{\partial \boldsymbol{\lambda}_i}$ and $\boldsymbol{ r} = (\mathcal{Y} - \mathcal{M})$. The entrie $(i,j)$ of the sensitivity matrix $Q \times Q$ of $\psi_{\boldsymbol{\lambda}}$ is given by

@@

\noindent The entrie $(i,j)$ of the variability matrix $Q \times Q$ of $\psi_{\boldsymbol{\lambda}}$ is defined by

@@

\noindent where $k_{l}^{(4)}$ denotes the fourth cumulative of $\mathcal{Y}_{l}$. In the McGLM estimation process, empirical versions are used.

@@

Let $\boldsymbol{\hat{\theta}} = (\boldsymbol{\hat{\beta}^{\top}}, \boldsymbol{\hat{\lambda}^{\top}})^{\top }$ the estimator based on estimation functions of $\boldsymbol{\theta}$. Then the asymptotic distribution of $\boldsymbol{\hat{\theta}}$ is

>>>>

The power variance function characterizes the Tweedie family of distributions, it is given by $\vartheta\left(\cdot; p_r\right) = \mu^{p_r}_r$, in which some distributions stand out: Normal ($p $ = 0), Poisson ($p$ = 1), gamma ($p$ = 2) and Inverse Gaussian ($p$ = 3) \cite{Jorgensen87, Jorgensen97}.

The Poisson–Tweedie dispersion function \cite{Jorgensen15} is an option for responses that characterize counts. The dispersion function is given by $\vartheta\left(\cdot; p\right) = \mu + \tau\mu^p$ where $\tau$ is the dispersion parameter. We thus have a rich class of models to deal with responses that characterize counts, since many important distributions appear as special cases, such as: Hermite ($p$ = 0), Neyman type A ($p$ = 1), negative binomial ($p$ = 2) and Poisson–inverse Gaussian (p = $3$).

@@

The dispersion matrix $\boldsymbol{\Omega({\tau})}$ describes the part of the covariance within each response variable that does not depend on the mean structure, ie, the correlation structure between the sample observations. Based on the ideas of \cite{Anderson73} and \cite{Pourahmadi00}, \cite{Bonat16} proposed to model the dispersion matrix through a matrix linear predictor combined with a covariance link function given by:

@@

In this way, the McGLMs configure a general framework for analysis via regression models for non-Gaussian data with multiple responses in which no assumptions are made regarding the independence of the observations. The class is defined by 3 functions (link, variance and covariance) in addition to a linear predictor and a matrix linear predictor for each response under analysis.

@@

McGLMs are fitted are fitted based on the estimating function approach described in detail by \cite{Bonat16} and \cite{jorg04}. This subsection presents an overview of the algorithm and the asymptotic distribution of estimators based on estimating functions.

@@

For the dispersion parameters, the Pearson estimating function is used, defined in the form

@@

\noindent where $W_{\boldsymbol{\lambda}i} = -\frac{\partial \boldsymbol{C}^{-1}}{\partial \boldsymbol{\lambda}_i}$ and $\boldsymbol{ r} = (\mathcal{Y} - \mathcal{M})$. The entry $(i,j)$ of the sensitivity matrix $Q \times Q$ of $\psi_{\boldsymbol{\lambda}}$ is given by

@@

\noindent The entry $(i,j)$ of the variability matrix $Q \times Q$ of $\psi_{\boldsymbol{\lambda}}$ is defined by

@@

\noindent where $k_{l}^{(4)}$ denotes the fourth cumulant of $\mathcal{Y}_{l}$. In the McGLM estimation process, empirical versions are used.

@@

Let $\boldsymbol{\hat{\theta}} = (\boldsymbol{\hat{\beta}^{\top}}, \boldsymbol{\hat{\lambda}^{\top}})^{\top }$ the estimator based on estimating functions of $\boldsymbol{\theta}$. Then the asymptotic distribution of $\boldsymbol{\hat{\theta}}$ is
